---
title: "The Wasserstein prior for the Skew Normal distribution"
author: '[F. Javier Rubio](https://sites.google.com/site/fjavierrubio67/)'
date: '`r format(Sys.Date(), "%B %d, %Y")`'
output:
  html_document:
    toc: true
    toc_float: true
bibliography: references.bib
---

# The Wasserstein prior for the Skew Normal distribution

The skew-normal pdf is defined as @azzalini:1985:
$$
s(x \mid \alpha) = 2 \phi(x) \Phi(\alpha x),
$$
where $\alpha \in {\mathbb R}$ is a skewness parameter.


@rubio:2022 calculated the The Wasserstein prior for the skewness parameter of the skew normal distribution 
$$
\pi_W(\alpha)  \propto \sqrt{\int_{-\infty}^{\infty} \frac{\sqrt{2} e^{-\frac{1}{2} \left(2 \alpha ^2+1\right) x^2}}{\pi ^{3/2} \left(\alpha
   ^2+1\right)^2 \left(\mbox{erf}\left(\frac{\alpha  x}{\sqrt{2}}\right)+1\right)}dx},
$$
They showed that this prior is symmetric about $0$, integrable, and has of order ${\mathcal O}(\vert \alpha \vert^{-5/2})$. They proposed approximating the Wasserstein prior using a Student-t distribution with $3/2$ degrees of freedom.

The following R code shows the shape of the Wasserstein prior and a comparisons with the proposed approximation.

See also: [Natural (non-) informative priors for the skew-normal distribution](https://rpubs.com/FJRubio/BTV) and [The Jeffreys prior for skew–symmetric models](https://rpubs.com/FJRubio/JPSS).

## The Wasserstein prior
```{r}
# Wasserstein information function
WIM <- Vectorize(function(alpha){
  # Integrand
  integ <- Vectorize(function(x){
    log_num <- 0.5*log(2) - 0.5*(1 + 2*alpha^2)*x^2
    log_den <- 1.5*log(pi) + 2*log(1 + alpha^2) - log(2) - pnorm(alpha*x, log.p = TRUE)
    out <- exp(log_num - log_den)
    return(out)
  })
  int <- integrate(integ,-Inf,Inf)$value
})

# Normalising constant
tempf <- Vectorize(function(alpha){sqrt(WIM(alpha))})
normc <- integrate(tempf,-Inf,Inf)$value

# Wasserstein prior
piW <- Vectorize(function(alpha){sqrt(WIM(alpha))/normc })

curve(piW,-5,5, lwd = 2, n = 1000, xlab = expression(alpha), ylab = "Prior density", main = "Wasserstein prior",
      cex.axis = 1.5, cex.lab = 1.5)
```

## Tractable approximation
```{r}
# t-approximation
sc = 0.757
appt <- Vectorize(function(x) dt(x/sc, df = 3/2)/sc)

curve(piW,-5,5, lwd = 2, n = 1000, xlab = expression(alpha), ylab = "Prior density", main = "Wasserstein prior",
      cex.axis = 1.5, cex.lab = 1.5)
curve(appt,-5,5, lwd = 2, lty = 2, add = T, n = 1000)
legend('topright', legend = c('Wasserstein prior', 't-approximation'), lwd = 2, lty = c(1,2))



abs_err <- Vectorize(function(x) abs(piW(x) - appt(x))  )

curve(abs_err,-5,5, lwd = 2, n = 1000, xlab = expression(alpha), ylab = "Absolute error", main = "",
      cex.axis = 1.5, cex.lab = 1.5)

```
